Effect of character correlation

### 2017-09-18 Analysis/EffectCorrelationStats.Rmd
Updated stats part

```{r, eval = TRUE, message = FALSE}
library(dispRity)
library(lme4)
library(lmerTest)
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")

@@

Summary statistics
-----------------------------------

@@

```{r}
## Running the pairwise Wilcox tests:
(sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL))
(sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL))
(sum_RF_null <- apply.test(whole_data, "RF", "null", NULL))
(sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL))

```

@@

## Summarise the results
results <- matrix(nrow = 4, data = c(
                round(unlist(lapply(normality, `[[`, 1)), digit = 3),
                round(unlist(lapply(normality, `[[`, 2)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 1)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 2)), digit = 3),
                round(unlist(lapply(homoscedasticity, `[[`, 3)), digit = 3)),
            dimnames = list(
                c("RF best", "Triples best", "RF null", "Triples null"),
                c("W", "p.value", "Bartlett's K-squared", "df", "p.value")))

## Check assumption's validities
validity <- apply(results, 1,
    function(X) ifelse((X[2] >= 0.05 && X[5] >= 0.05), TRUE, FALSE))

## Report results
(results <- cbind(as.data.frame(results), validity))

```

@@

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don't overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
## Running the pairwise Bhattacharrya Coefficient analysis:
(bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff))
(bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff))
(bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff))
(bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff))
```

@@

```{r}
## Running the pairwise Wilcox tests:
wx_RF_best <- apply.test(whole_data, "RF", "norm", wilcox.test)
wx_Tr_best <- apply.test(whole_data, "Triples", "norm", wilcox.test)
wx_RF_null <- apply.test(whole_data, "RF", "null", wilcox.test)
wx_Tr_null <- apply.test(whole_data, "Triples", "null", wilcox.test)

## Applying Bonferroni correction
wx_RF_best[,c(2,4,6)] <- p.adjust(wx_RF_best[,c(2,4,6)], method = "bonferroni")
wx_Tr_best[,c(2,4,6)] <- p.adjust(wx_Tr_best[,c(2,4,6)], method = "bonferroni")
wx_RF_null[,c(2,4,6)] <- p.adjust(wx_RF_null[,c(2,4,6)], method = "bonferroni")
wx_Tr_null[,c(2,4,6)] <- p.adjust(wx_Tr_null[,c(2,4,6)], method = "bonferroni")

## Print the results
wx_RF_best
wx_Tr_best
wx_RF_null
wx_Tr_null
```

>>>>

```{r, eval = TRUE, message = FALSE}
library(knitr)
library(dispRity)
library(lme4)
library(lmerTest)
source("functions.R") ; load.functions(test = FALSE)
dyn.load("../Functions/char.diff.so")
load("../Data/NTS/matrix.nts.Rda")
source("../Data/load.data.R")

@@

Pooled results
=================

Individual results
=================

We can also analyse the effect of each parameters combination separately:

Summary statistics
-----------------

@@

```{r}
## Running the pairwise Wilcox tests:
sum_RF_best <- apply.test(whole_data, "RF", "norm", NULL)
sum_Tr_best <- apply.test(whole_data, "Triples", "norm", NULL)
sum_RF_null <- apply.test(whole_data, "RF", "null", NULL)
sum_Tr_null <- apply.test(whole_data, "Triples", "null", NULL)
```

```{r, echo = FALSE}
## Setting the table captions
captions_summary <- paste("Summary statistics of the normalised", rep(c("Robinson-Fould", "Triplets"), 2), "distance to the", c(rep("best", 2), rep("null", 2)), "tree.")

## Printing the tables
kable(sum_RF_best, row.names = FALSE, caption = captions_summary[1])
kable(sum_Tr_best, row.names = FALSE, caption = captions_summary[2])
kable(sum_RF_null, row.names = FALSE, caption = captions_summary[3])
kable(sum_Tr_null, row.names = FALSE, caption = captions_summary[4])
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
library(xtable)
## Wrapper function
make.xtable <- function(table, digit = 3, caption, label, longtable = FALSE) {

  table <- xtable(table, digit = digit, caption = caption, label = label)

  if(longtable == TRUE) {
    print(table, tabular.environment = 'longtable', floating = FALSE, include.rownames = FALSE)
  } else {
    print(table, include.rownames = FALSE)
  }
}

## Labels
labels_summary <- paste0("Tab_Summary", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

## Tables
make.xtable(sum_RF_best, caption = captions_summary[1], label = labels_summary[1])
make.xtable(sum_Tr_best, caption = captions_summary[2], label = labels_summary[2])
make.xtable(sum_RF_null, caption = captions_summary[3], label = labels_summary[3])
make.xtable(sum_Tr_null, caption = captions_summary[4], label = labels_summary[4])
```

@@

# ## Summarise the results
results <- t(rbind(sapply(normality, function(x) x[c(1,2)]), 
                   sapply(homoscedasticity, function(x) x[c(1,2,3)])))
results <- matrix(as.numeric(results), ncol = 5)
colnames(results) <- c("W", "p.value", "Bartlett's K^2", "df", "p.value")
rownames(results) <- c("RF best", "Triples best", "RF null", "Triples null")
results <- round(results, 3)

# ## Check assumption's validities
check.validity <- function(X) {ifelse(X[2] >= 0.05 && X[5] >= 0.05, TRUE, FALSE)}
validity <- apply(results, 1, check.validity)

## Report results
results <- cbind(as.data.frame(results), validity)
```

```{r, echo = FALSE}
kable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 3)
```

<!-- Table for publication -->

```{r, eval = FALSE, echo = FALSE}
xtable(results, caption = "Normality and variance homoscedasticity tests for anova.", digit = 2, label = "Tab_signif")
```

@@

We can also look at the effect of each scenario within each parameter setting (e.g. 25 taxa, 100 characters and Bayesian trees) by measuring the probability of overlap between both distributions.
A probability of overlap probability below 0.05 means the two scenario gives a distribution of tree scores that don t overlap (significant effect of one scenario on topology).
Conversely, an overlap probability above 0.95 means the two scenario gives the same tree scores (significant no effect of one scenario on topology).

```{r}
## Running the pairwise Bhattacharrya Coefficient analysis:
bc_RF_best <- apply.test(whole_data, "RF", "norm", bhatt.coeff)
bc_Tr_best <- apply.test(whole_data, "Triples", "norm", bhatt.coeff)
bc_RF_null <- apply.test(whole_data, "RF", "null", bhatt.coeff)
bc_Tr_null <- apply.test(whole_data, "Triples", "null", bhatt.coeff)

## Tables
kable(bc_RF_best, row.names = FALSE, digit = 3)
kable(bc_RF_best, row.names = FALSE, digit = 3)
kable(bc_RF_best, row.names = FALSE, digit = 3)
kable(bc_RF_best, row.names = FALSE, digit = 3)
```

```{r, echo = FALSE}
## Setting the table captions
captions_bc <- paste0("Bhattacharrya Coefficient (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios.")

## Printing the tables
kable(bc_RF_best, row.names = FALSE, caption = captions_bc[1], digit = 3)
kable(bc_Tr_best, row.names = FALSE, caption = captions_bc[2], digit = 3)
kable(bc_RF_null, row.names = FALSE, caption = captions_bc[3], digit = 3)
kable(bc_Tr_null, row.names = FALSE, caption = captions_bc[4], digit = 3)
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_bc <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

## Tables
make.xtable(bc_RF_best, caption = captions_bc[1], label = labels_bc[1])
make.xtable(bc_Tr_best, caption = captions_bc[2], label = labels_bc[2])
make.xtable(bc_RF_null, caption = captions_bc[3], label = labels_bc[3])
make.xtable(bc_Tr_null, caption = captions_bc[4], label = labels_bc[4])
```

@@

```{r, warning = FALSE}
## Running the pairwise Wilcox tests:
wx_RF_best <- apply.test(whole_data, "RF", "norm", wilcox.test)
wx_Tr_best <- apply.test(whole_data, "Triples", "norm", wilcox.test)
wx_RF_null <- apply.test(whole_data, "RF", "null", wilcox.test)
wx_Tr_null <- apply.test(whole_data, "Triples", "null", wilcox.test)

## Applying Bonferroni correction
for(column in c(5,7,9)) {
  wx_RF_best[,column] <- p.adjust(wx_RF_best[,column], method = "bonferroni")
  wx_Tr_best[,column] <- p.adjust(wx_Tr_best[,column], method = "bonferroni")
  wx_RF_null[,column] <- p.adjust(wx_RF_null[,column], method = "bonferroni")
  wx_Tr_null[,column] <- p.adjust(wx_Tr_null[,column], method = "bonferroni")
}
```

```{r, echo = FALSE}
## Setting the table captions
captions_wx <- paste0("Non-parametric Wilcoxon test (",rep(c("Robinson-Fould distance", "Triplets distance"), 2) ," from the ",c(rep("best", 2), rep("null", 2)), " tree)" , " between the different scenarios (p-values corrected using Bonferroni-Holm correction).")

## Printing the tables
kable(wx_RF_best, row.names = FALSE, caption = captions_wx[1], digit = 3)
kable(wx_Tr_best, row.names = FALSE, caption = captions_wx[2], digit = 3)
kable(wx_RF_null, row.names = FALSE, caption = captions_wx[3], digit = 3)
kable(wx_Tr_null, row.names = FALSE, caption = captions_wx[4], digit = 3)
```

<!-- Table for publication -->

```{r, echo = FALSE, eval = FALSE}
## Labels
labels_wx <- paste0("Tab_BC", rep(c("RF", "Tr"), 2), c(rep("best", 2), rep("null", 2)))

## Tables
make.xtable(wx_RF_best, caption = captions_wx[1], label = labels_wx[1])
make.xtable(wx_Tr_best, caption = captions_wx[2], label = labels_wx[2])
make.xtable(wx_RF_null, caption = captions_wx[3], label = labels_wx[3])
make.xtable(wx_Tr_null, caption = captions_wx[4], label = labels_wx[4])
```
